Brownian motion of main-belt asteroids on human timescales

### 2025-08-19 response.txt

Response to first referee report.

I thank the referee for their concise and insightful review of the paper.  The
main issue raised is that this paper relies heavily on the companion paper
analysing LSST constraints on asteroid masses from mutual encounters.  That
paper was submitted to PSJ some weeks before this one, and has since been
refereed, revised, accepted, and published in the journal, which I hope will
assuage this worry.

One of the revisions made to that paper addressed an issue also raised by this
referee, which was the rather high value of albedo assumed for the main-belt
asteroid population.  To address this concern for the other paper, we changed
our conversion from (well-observed) H values to (physically relevant) mass
substantially,  Instead of choosing a representative value of  both density and
albedo of the population, we surveyed the literature for individual main-belt
asteroids (MBAs) with mass determinations of any useful precision, then examined
the ratio $M/ 10^{-0.6H}$ directly.  The number of such bodies is actually
larger than the number for which both density and albedo have been determined
to similar precision, so is a simpler and slightly more accurate means of deriving
the "mass-to-light" conversion.  See Figure 1 and Section 2.1 of the published
paper.

The adopted result was that the distribution of mass-to-light
ratios of this (incomplete, heterogeneous, almost certainly biased) population
could be crudely characterized by taking a "high" and "low" value, which differ
by a factor of 5.  The main change made to this Brownian-motion paper is to
adopt these two values and recalculate the resulting integrals that yield the
estimated RMS 10-year uncertainties that are the key result of this paper.
These mass-to-light ratio changes appear as multiplicative scalings of the
previous result.  The thresholds for individually-detectable impulses in Section
3.2.3 have also been updated in concert with the updates to the
mass-determination paper.

Thus the numerical results presented in this paper have been adjusted, but as
per the referee's comments, the methodology remains unchanged.  Changed text is
in red in the draft, and figures that have changed have this noted in their
captions.

A few minor replies to the referee's comments are below.

> This report evaluates the manuscript "Brownian Motion of Main-Belt
> Asteroids on Human Timescales," which examines how gravitational
> interactions between asteroids introduce stochastic positional
> uncertainties over decadal timescales. The study provides a novel
> approach to understanding perturbative effects in asteroid
> ephemerides, with implications for high-precision astronomy.
> 
> The central hypothesis of the study is that the long-term orbital
> evolution of main-belt asteroids is affected by random gravitational
> impulses from other asteroids, leading to a Brownian motion-like
> positional drift. This effect is expected to be particularly
> significant over human timescales (10-100 years), impacting asteroid
> tracking and planetary defense efforts. The study proposes that this
> motion is primarily azimuthal and that current ephemeris models
> underestimate this stochastic contribution.

We did not intend to give the impression that Brownian motion is larger or more
important on decadal scales than over millennia.  Rather our purpose is to
quantify the effect of the Brownian motion on "real-time" dynamics of the solar
system, i.e. the development of ephemerides that fit obervations and adjust
parameters of the solar system's members and of the laws of gravitation.
Understanding the long-term dynamical consequences of the Brownian motion for
population evolution is beyond our scope.

> 
> The study is based on several key assumptions:
> 
> - Asteroid Encounters as Stochastic Events: The model assumes that
> gravitational encounters between asteroids can be treated as a series
> of independent random impulses.
> 
> - Negligibility of Non-Gravitational Effects: The study assumes that
> radiation forces (e.g., Yarkovsky effect) and thermal recoil forces
> have a minor influence compared to gravitational interactions for
> asteroids larger than ~100 m.
> 
> - Mass Uncertainty from LSST Data: The study uses projected LSST data
> to estimate mass uncertainties, heavily relying on Bernstein et
> al. (2025) for these calculations.
> 
> - Choice of Geometric Albedo: The author chooses the geometric albedo
> value that might, in this evaluator's judgment, be too "bright" to be
> representative of the entire population of asteroids in the main
> belt. The problem with that choice lies in the fact that, with that
> value, the mass distribution estimate would be misrepresented. A value
> of 0.15 might be more representative of the average, so I recommend
> its use.
>

See the note at top for a major revision to our estimation of asteroid masses
from their absolute magnitudes.

> In terms of methodology, the study employs a combination of
> theoretical modeling, statistical analysis, and observational
> constraints:
> 
> -Impulse-Based Orbital Perturbation Model: The author developed a
> model where asteroid encounters introduce small, random velocity
> perturbations that accumulate over time, following a Brownian motion
> framework.
> 
> -Empirical Scaling Relations: The study derives scaling relations for
> the expected positional variance over time.
> 
> -Comparison with Gaia and LSST Predictions: The paper estimates how
> well future astronomical surveys can track and mitigate these
> stochastic uncertainties.  Both the mathematical approach and its
> development are appropriate, and I have not detected any errors in
> them.
> 
> Regarding the results, the manuscript highlights some points that
> might be very interesting (the reasons for using "might" are explained
> below):
> 
> - the RMS azimuthal shift due to Brownian motion reaches ~2 km over 10
> years (~1 mas in angular uncertainty).
> 
> - Radial and vertical displacements are found to be significantly
> smaller, with the azimuthal motion dominating the effect.
> 
> - LSST mass uncertainty reductions could improve orbit predictions but
> do not eliminate Brownian motion noise.
> 
> Although the study presents an interesting and underexplored
> perspective on stochastic effects affecting asteroid motion and has a
> robust framework with an application of "Brownian motion" to celestial
> mechanics, it is fully justified to rely on a paper that is in
> preparation (Bernstein et al., 2025) and has clearly not gone through
> the peer review process. This is the reason for using "might" in the
> results section, making it difficult to analyze the final results. If
> the author can improve on this point, the paper would be an
> interesting contribution to the field.
>

### 2025-08-20 main.tex
Submission 2

\newcommand{\ra}{\mathrm{RA}}
\newcommand{\dec}{\mathrm{Dec}}
\newcommand{\lsst}{\textit{LSST}}
\newcommand{\gaia}{\textit{Gaia}}
\newcommand{\AU}{\mathrm{au}}
\newcommand{\eqq}[1]{Equation~(\ref{#1})}
\newcommand{\ie}{\textit{i.e.\/}}
\newcommand{\eg}{\textit{e.g.\/}}
\newcommand\edited[1]{{\color{blue} {#1}}}
\newcommand\gary[1]{{\color{red} {\textbf{GMB}: #1}}}
% Turn off change highlighting
%\newcommand\edited[1]{#1}

@@

\newcommand{\Var}{\textrm{Var}}
\newcommand{\vcirc}{v_c}
\newcommand{\Msun}{M_\odot}
\newcommand{\covm}{C}
\newcommand{\lop}{\varpi}   % longitude of perihelion
\newcommand{\ma}{\ell}  % Mean anomaly

@@

\begin{abstract}
  % The explosion of high-precision astrometric data on main-belt
  % asteroids (MBAs) enables new
  % inferences on gravitational and non-gravitational forces present in
  % this region.  We estimate the size of MBA motions
  % caused by mutual gravitational encounters with other MBAs that are
  % either omitted from ephemeris models or have uncertain mass
  % estimates. In other words, what is the typical Brownian motion among
  % MBAs that cannot be predicted from the ephemeris, and therefore
  % serves as noise on inferences from the MBAs?  We estimate the RMS
  % azimuthal shift $\sigma_\phi$ of this ``Brownian noise''  by
  % numerical estimation of the distribution of impulse sizes and
  % directions among known MBAs, combined with analytical propagation
  % into future positional uncertainties. At current
  % levels of asteroid-mass knowledge, $\sigma_\phi$ rises to
  % $\approx2$~km or $\approx1$~mas over $T=10$~yr, increasing as
  % $T^{3/2},$  large enough to degrade many inferences from
  % Gaia and LSST MBA data. LSST data will, however, improve MBA
  % mass knowledge enough to lower this Brownian uncertainty by
  % $\ge4\times.$  Radial and vertical Brownian noise  at $T=10$~yr are
  % factors $\approx7$ and $\approx45,$ respectively,
  % lower than the azimuthal noise, and grow as $T^{3/2}$
  % and $T^{1/2}.$  For full exploitation of Gaia and
  % LSST MBA data, ephemeris models should include the
  % $\approx1000$~largest asteroids as active bodies with free masses,
  % even if not all are well constrained.  This will correctly propagate
  % the uncertainties from these 1000 sources' deflections into desired
  % inferences. The RMS value of deflections from less-massive MBAs is
  % then just $\sigma_\phi\approx60$~m or 30~$\mu$as, small enough to
  % ignore until occultation-based position data become ubiquitously
  % available for MBAs.
%
High-precision astrometric data on main-belt
asteroids (MBAs) enables new
inferences on gravitational and non-gravitational forces present in
this region.  We estimate the typical Brownian motion due to mutual encounters among
MBAs that cannot be predicted from the ephemeris of larger bodies, and therefore
serves as noise on inferences from the MBAs.  The RMS
azimuthal shift $\sigma_\phi$ of this ``Brownian noise''  is predicted
using the distribution of known MBA orbits, combined with analytical propagation
into future positional uncertainties. At current
levels of asteroid-mass knowledge, $\sigma_\phi$ rises to
$\approx2$~km or $\approx1$~mas over $T=10$~yr, increasing as
$T^{3/2},$  large enough to degrade many inferences from
Gaia and LSST MBA data. LSST data will, however, improve MBA
mass knowledge enough to lower this Brownian uncertainty by
$\ge4\times.$  Radial (vertical) Brownian noise  at $T=10$~yr is
$\approx7\times$ ($\approx45\times$) 
lower than the azimuthal noise.  For full exploitation of Gaia and
LSST MBA data, ephemeris models should include the
$\approx1000$~largest asteroids as active bodies with free masses,
even if not all are well constrained.  This will correctly propagate
the uncertainties from these 1000 sources' deflections into desired
inferences. The RMS value of deflections from less-massive MBAs is
then just $\sigma_\phi\approx60$~m or 30~$\mu$as, small enough to ignore until occultation-based position data become ubiquitously available for MBAs.
\end{abstract}

@@

High-precision measurements of positions of Solar System bodies have
great value to multiple scientific pursuits.  They can reveal
gravitational accelerations due to undiscovered bodies such as a
hypothetical ``Planet X'' \citep[\eg][]{holmanP9,inpopP9,trojans,occultations} or passing primordial
black holes \citep[\eg][]{pbh}, determine
the masses of known bodies \citep[\eg][]{goffin,baer}, test the laws
of physics \citep[\eg][]{inpopgraviton}, and characterize 
the non-gravitational forces that drive many forms of planetary
migration, such as the maintainence of the near-Earth asteroid (NEA)
population \citep[reviewed by][]{yarkovsky}.  While most applications of precision orbit determination
have used the 8 major planets as test particles, there are now $>10^6$
known minor planets available for this task.  These small bodies' astrometric
information content is growing extremely rapidly through large-scale
survey projects that can produce
milliarcsecond-scale accuracy for magnitude-limited small-body populations, such as Gaia \citep{gaiass3,gaiafpr},
PanSTARRS,\footnote{\url{https://www2.ifa.hawaii.edu/research/Pan-STARRS.shtml}}
and particularly the upcoming Rubin observatory
\textit{Legacy Survey of Space and Time} (LSST).\footnote{\url{https://rubinobservatory.org/}}  The great majority
of currently-tracked objects are main-belt asteroids (MBAs), so we
wish to investigate the potential power of MBAs as gravitational test
bodies.

@@

\section{Calculation overview}
We aim to calculate the Brownian motion variance to an accuracy of a
factor $\approx 2.$
Greater accuracy in the calculation is not warranted
since some of the principal inputs have substantial uncertainty.  The
number of MBAs vs mass is 
uncertain because of unknowns in the conversion
from $H$ to mass; the status of future observations and $\sigma_M$ is
also unclear.  We are hence justified in retaining only the
contributions to our results that are leading
order in the orbital eccentricity $e$ of the tracer MBA.
We will assume throughout that the joint
distribution of mass (or $H$) and orbital elements is separable into a
mass distribution and an orbital distribution.  In this scenario, the
Brownian motion is independent of the tracer mass, and involves an
integral over the mass distribution of the deflecting bodies.

We adopt a cylindrical coordinate system $(r,\phi,z),$ with $\zhat$
normal to the initial orbital plane, and $\phi=z=0$ toward the
perhelion of the initial orbit, \ie\ both the initial ascending node
$\Omega_0$ and longitude of perihelion $\lop_0$ equal to zero.  The
unit vectors $\rhat, \phat$ rotate with the target asteroid.  The mean
anomaly at time $t=0$ is $\ma_0.$  In general, subscripts of $0$ will
indicate properties of the unperturbed orbit.

@@

\section{Impulse distribution}
\label{sec:impulse}
\subsection{Derivation}
Under the assumption that the distributions of masses and orbital
elements of MBAs are separable, the quantities needed to describe the
random walk of the asteroid's orbits can be expressed as an integral
over the absolute magnitude $H$ of the deflectors, the relative velocity $v$,
impact parameter $b,$ and polar/azimuthal angles $\theta,\phi$ of the
unit vector of the impulse direction $\bhat$:
\begin{eqnarray}
  \left\langle n I_\gamma^2 \right\rangle & = & \int
                                              dH\,dv\,db\,d\theta\,d\phi
                                             \frac{dN}{dH} \frac{dn}{dv\,db\,d\theta\,d\phi}
                                              I^2_\gamma(H,v,b,\theta,\phi)
  \\
  & = & \int dH \frac{dN}{dH} \int dv
        \frac{dn}{dA\,dv\,d\theta\,d\phi} \hat b_\gamma^2(\theta,\phi)
        \int_{b_{\rm min}}^{b_{\rm max}} 2\pi b\,db\,
        \left(\frac{2\sigma_M(H)}{bv}\right)^2 \\
  & = & 8\pi \int dH \frac{dN}{dH} \sigma^2_M(H)  \int dv
        \frac{dn}{dA\,dv} v^{-2} \left\langle\hat b_\gamma^2(v)\right\rangle
        \log(b_{\rm max}/b_{\rm min}).
\end{eqnarray}
The second row adopts the (numerically verified) assumption that the
distribution of impact parameter $b$ will always be $\propto dA=2\pi
b\,db$ in the range of $b$ of interest to us.  The mean rate of encounters
between a deflector-tracer pair, as a function of the impact velocity
and geometry, is $dn/dA\,dv\,d\theta\,d\phi,$ which we will determine
through numerical measurements with the orbits of all known MBAs. The
$H$ distribution of MBAs is $dN/dH.$ We define $\sigma_M(H)$ via
\begin{equation}
  \sigma_M(H) \equiv \left\{\begin{array}{cl}
                              \sigma_M, & M(H) > \sigma_M \\
                              M(H),        & M(H) \le \sigma_M
                            \end{array} \right.
  \label{eq:sigmaMH}
\end{equation}
such that it represents the error in the ephemeris due to uncertainty
in the masses of MBAs included in the ephemeris (first row), and due
to the entirety of the masses of MBAs omitted from the ephemeris
(second row).

In the third line, we execute the integral over $b$ and introduce the
average geometry factors $\langle \hat b^2_\gamma\rangle$ of encounters as
a function of $v$.  It must be true that
\begin{equation}
  \left\langle \hat b_r^2 \right\rangle
  + \left\langle \hat b_\phi^2 \right\rangle
  + \left\langle \hat b_z^2 \right\rangle = 1,
\end{equation}
but equality of the three is not required.

@@

\subsection{Numerical results}
\eqq{eq:nI2} reduces the calculation to a double integral over the
deflector $H$ and the tracer-deflector relative velocity $v.$  The
constituents of this calculation are estimated as follows.

@@

\subsubsection{Asteroid properties}
\label{sec:mbaprops}
We crudely approximate the population as having a geometric albedo of 0.25,
density of 2500~kg~m$^{-3},$ and spherical shapes, which leads to the
conversion
\begin{equation}
  M(H) = 1.2\times10^{-17} M_\odot \times 10^{-0.6(H-15)}
\end{equation}

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{mbacounts.pdf}
  \caption{Differential counts of MBAs cataloged by the MPC are
    plotted vs absolute magnitude $H$  
    for the Inner, Middle, Outer, and ``Other'' regions of
    the main belt, and for their total.  We assume that they are
    complete for $H\le17,$ and extrapolate to fainter sources using
    the functional form from \eqq{eq:lsstmba} shown as the dotted
    curve. The Inner and Outer regions are slightly rescaled to help
    visual comparison of the curves' shapes in the three MBA regions.}
  \label{fig:counts}
\end{figure}
We also require the $H$ distribution of potential deflectors.
Figure~\ref{fig:counts} plots our assumptions for $dN/dH.$  The MPC
catalog is likely close to complete in all three regions for $H\le17,$
so we will use the MPC counts directly in this regime.  Note that the
three regions have similar shapes of $dN/dH.$  For $H>17,$ we adopt
the functional form for the cumulative MBA counts given by
initial LSST predictions \citep{lsstbook}:
\begin{equation}
  N(<H) \propto \frac{10^{0.43(H-15.7)}}{10^{0.18(H-15.7)} +
    10^{-0.18(H-15.7)}},
\label{eq:lsstmba}
\end{equation}
normalizing this curve to the observed $H<17$ counts for each region.

Our integral for the uncertainty due to unmodelled MBA gravitation
requires the RMS error $\sigma_M(H)$ between the true mass $M$ and the
mass used (if any) in the ephemeris model.
The two largest asteroids, Ceres and Vesta, have masses determined to
high precision by the \textit{Dawn} spacecraft \citep{Ceres,Vesta}.  Thus despite holding
half of the $\approx 10^{-9} M_\odot$ mass of the asteroid belt, their $\sigma_M$
values\footnote{More precisely, the uncertainties in their $GM$ values
  relative to $GM_\odot$.} are 
$\le3\times10^{-15} M_\odot$ and
% Konilov papers: 1.2e-5 km^3/s^2 for Vesta, 4e-4 for Ceres
this unknown part of their contribution to other MBA's motion is 
unimportant.  Aside from a handful of other asteroids in
well-characterized binary systems or visited by spacecraft, most other knowledge of MBA masses
has and will come from mutual gravitational encounters.   An
individual asteroid's mass can be estimated from its measured effect  on the
few MBAs on which it imparts the largest impulses.  The accuracy
$\sigma_M$ of this determination will vary, depending upon the
geometry  and timing of
each deflector MBA's encounters, and the measurement errors on its tracers.  These
factors are independent of the mass of the deflector.  We will
assign a value $\sigma_M$ to the RMS uncertainty in MBA mass
attainable from following a few individual tracers, and calculate
the Brownian-motion noise as a function of $\sigma_M.$

\begin{figure}
  \plottwo{sigmaI.pdf}{sigmaM.pdf}
  \caption{Forecasts of uncertainties derived from LSST tracking of the known MBAs.  The left-hand plot shows the distribution of (log) uncertainties $\sigma_I$ on the amplitude of the impulse applied to the tracer MBA, incorporating forecasted LSST observations and marginalizing over the tracer's initial state vector.  The right-hand side shows the distribution of the uncertainties on the mass, $\sigma_M=(bv/2)\sigma_I,$ for deflector MBAs expected from LSST.  The histograms show the results from using only the single tracer yielding the lowest $\sigma_M$ for that deflector; combining the 10 best-$\sigma_M$ tracers; or combining all impulses at $b<0.03$~AU.  The RMS value of $\sigma_M$ over the deflector population is given in the legend.  Masses are in solar units, and velocities in units of the circular velocity of the deflector.}
  \label{fig:sigmaI}
\end{figure}

>>>>

\newcommand{\ra}{\mathrm{RA}}
\newcommand{\dec}{\mathrm{Dec}}
\newcommand{\lsst}{\textit{LSST}}
\newcommand{\gaia}{\textit{Gaia}}
\newcommand{\AU}{\mathrm{au}}
\newcommand{\eqq}[1]{Equation~(\ref{#1})}
\newcommand{\ie}{\textit{i.e.\/}}
\newcommand{\eg}{\textit{e.g.\/}}
\newcommand\edited[1]{{\color{red} {#1}}}
\newcommand\gary[1]{{\color{red} {\textbf{GMB}: #1}}}
% Turn off change highlighting
%\newcommand\edited[1]{#1}

@@

\newcommand{\Var}{\textrm{Var}}
\newcommand{\vcirc}{v_c}
\newcommand{\Msun}{M_\odot}
\newcommand{\covm}{C}
\newcommand{\lop}{\varpi}   % longitude of perihelion
\newcommand{\ma}{\ell}  % Mean anomaly
\newcommand{\uas}{$\mu$as}

@@

\begin{abstract}
  % The explosion of high-precision astrometric data on main-belt
  % asteroids (MBAs) enables new
  % inferences on gravitational and non-gravitational forces present in
  % this region.  We estimate the size of MBA motions
  % caused by mutual gravitational encounters with other MBAs that are
  % either omitted from ephemeris models or have uncertain mass
  % estimates. In other words, what is the typical Brownian motion among
  % MBAs that cannot be predicted from the ephemeris, and therefore
  % serves as noise on inferences from the MBAs?  We estimate the RMS
  % azimuthal shift $\sigma_\phi$ of this ``Brownian noise''  by
  % numerical estimation of the distribution of impulse sizes and
  % directions among known MBAs, combined with analytical propagation
  % into future positional uncertainties. At current
  % levels of asteroid-mass knowledge, $\sigma_\phi$ rises to
  % $\approx2$~km or $\approx1$~mas over $T=10$~yr, increasing as
  % $T^{3/2},$  large enough to degrade many inferences from
  % Gaia and LSST MBA data. LSST data will, however, improve MBA
  % mass knowledge enough to lower this Brownian uncertainty by
  % $\ge4\times.$  Radial and vertical Brownian noise  at $T=10$~yr are
  % factors $\approx7$ and $\approx45,$ respectively,
  % lower than the azimuthal noise, and grow as $T^{3/2}$
  % and $T^{1/2}.$  For full exploitation of Gaia and
  % LSST MBA data, ephemeris models should include the
  % $\approx1000$~largest asteroids as active bodies with free masses,
  % even if not all are well constrained.  This will correctly propagate
  % the uncertainties from these 1000 sources' deflections into desired
  % inferences. The RMS value of deflections from less-massive MBAs is
  % then just $\sigma_\phi\approx60$~m or 30~$\mu$as, small enough to
  % ignore until occultation-based position data become ubiquitously
  % available for MBAs.
%
High-precision astrometric data on main-belt
asteroids (MBAs) enables new
inferences on gravitational and non-gravitational forces present in
this region.  We estimate the typical Brownian motion due to mutual encounters among
MBAs that cannot be predicted from the ephemeris of larger bodies, and therefore
serves as noise on inferences from the MBAs.  The RMS
azimuthal shift $\sigma_\phi$ of this ``Brownian noise''  is predicted
using the distribution of known MBA orbits, combined with analytical propagation
into future positional uncertainties---radial and vertical perturbations are
much smaller. At current
levels of asteroid-mass knowledge, \edited{we estimate $\sigma_\phi$ accumulates
to 2--9~km, or 1--5~mas over $T=10$~yr,  increasing as
$T^{3/2}.$  This must be taken into consideration for proper inferences from Gaia astrometry.}
LSST data will, however, improve MBA
mass knowledge enough to lower this Brownian uncertainty \edited{to 260--400~m,
  or 140--210~\uas, which in many but not all cases could be small enough to
  ignore in inferences using Gaia or LSST astrometry of MBAs.}
% Radial (vertical) Brownian noise  at $T=10$~yr is
% $\approx7\times$ ($\approx45\times$) 
% lower than the azimuthal noise.
For full exploitation of Gaia and
LSST MBA data, ephemeris models should include the
\edited{$\approx10,000$ largest} asteroids as active bodies with free masses,
even if not all are well constrained.  This will correctly propagate
the uncertainties \edited{in these bodies' gravity} into desired
inferences. The RMS value of deflections from less-massive MBAs is
then just \edited{$\sigma_\phi\approx17$--60~m, or 9--40~\uas, depending on the
  mean mass of MBAs,} small enough to ignore until occultation-based position
data become ubiquitously available for MBAs. 
\end{abstract}

@@

High-precision measurements of positions of Solar System bodies have
great value to multiple scientific pursuits.  They can reveal
gravitational accelerations due to undiscovered bodies such as a
hypothetical ``Planet X'' \citep[\eg][]{holmanP9,inpopP9,trojans,occultations} or passing primordial
black holes \citep[\eg][]{pbh}, determine
the masses of known bodies \citep[\eg][]{goffin,baer}, test the laws
of physics \citep[\eg][]{inpopgraviton}, and characterize 
the non-gravitational forces that drive many forms of planetary
migration, such as the maintenance of the near-Earth asteroid (NEA)
population \citep[reviewed by][]{yarkovsky}.  While most applications of precision orbit determination
have used the 8 major planets as test particles, there are now $>10^6$
known minor planets available for this task.  These small bodies' astrometric
information content is growing extremely rapidly through large-scale
survey projects that can produce
milliarcsecond-scale accuracy for magnitude-limited small-body populations, such as Gaia \citep{gaiass3,gaiafpr},
PanSTARRS,\footnote{\url{https://www2.ifa.hawaii.edu/research/Pan-STARRS.shtml}}
and particularly the upcoming Rubin observatory
\textit{Legacy Survey of Space and Time} (LSST).\footnote{\url{https://rubinobservatory.org/}}  The great majority
of currently-tracked objects are main-belt asteroids (MBAs), so we
wish to investigate the potential power of MBAs as gravitational test
bodies.

@@

\section{Calculation overview}
We aim to derive the Brownian motion variance to an accuracy of a
factor $\approx 2.$
Greater accuracy in the \edited{methodology is not warranted
since some of the principal inputs have substantial uncertainty, \eg\ the}
number of MBAs vs mass is 
uncertain because of unknowns in the conversion
from $H$ to mass, and the status of future observations and $\sigma_M$ is
also unclear.  We are hence justified in retaining only the
contributions to our results that are leading
order in the orbital eccentricity $e$ of the tracer MBA.
We will assume throughout that the joint
distribution of mass (or $H$) and orbital elements is separable into a
mass distribution and an orbital distribution.  In this scenario, the
Brownian motion is independent of the tracer mass, and involves an
integral over the mass distribution of the deflecting bodies.

We adopt a cylindrical coordinate system $(r,\phi,z),$ with $\zhat$
normal to the initial orbital plane, and $\phi=z=0$ toward the
perihelion of the initial orbit, \ie\ both the initial ascending node
$\Omega_0$ and longitude of perihelion $\lop_0$ equal to zero.  The
unit vectors $\rhat, \phat$ rotate with the target asteroid.  The mean
anomaly at time $t=0$ is $\ma_0.$  In general, subscripts of $0$ will
indicate properties of the unperturbed orbit.

@@

\section{Impulse distribution}
\label{sec:impulse}
\subsection{Derivation}
Under the assumption that the distributions of masses and orbital
elements of MBAs are separable, the quantities needed to describe the
random walk of the asteroid's orbits can be expressed as an integral
over the absolute magnitude $H$ of the deflectors, the relative velocity $v$,
impact parameter $b,$ and polar/azimuthal angles $\theta,\phi$ of the
unit vector of the impulse direction $\bhat$:
\begin{eqnarray}
  \left\langle n I_\gamma^2 \right\rangle & = & \int
                                              dH\,dv\,db\,d\theta\,d\phi
                                             \frac{dN}{dH} \frac{dn}{dv\,db\,d\theta\,d\phi}
                                              I^2_\gamma(H,v,b,\theta,\phi)
  \\
  & = & \int dH \frac{dN}{dH} \int dv\,\edited{d\theta\,d\phi}
        \frac{dn}{dA\,dv\,d\theta\,d\phi} \hat b_\gamma^2(\theta,\phi)
        \int_{b_{\rm min}}^{b_{\rm max}} 2\pi b\,db\,
        \left(\frac{2\sigma_M(H)}{bv}\right)^2 \\
  & = & 8\pi \int dH \frac{dN}{dH} \sigma^2_M(H)  \int dv
        \frac{dn}{dA\,dv} v^{-2} \left\langle\hat b_\gamma^2(v)\right\rangle
        \log(b_{\rm max}/b_{\rm min}).
\end{eqnarray}
The second row adopts the (numerically verified) assumption that the
distribution of impact parameter $b$ will always be $\propto dA=2\pi
b\,db$ in the range of $b$ of interest to us.  The mean rate of encounters
between a deflector-tracer pair, as a function of the impact velocity
and geometry, is $dn/dA\,dv\,d\theta\,d\phi,$ which we will determine
through numerical measurements with the orbits of all known MBAs. The
$H$ distribution of MBAs is $dN/dH.$ We define $\sigma_M(H)$ via
\begin{equation}
  \sigma_M(H) \equiv \left\{\begin{array}{cl}
                              \sigma_M, & M(H) > \sigma_M \\
                              M(H),        & M(H) \le \sigma_M
                            \end{array} \right.
  \label{eq:sigmaMH}
\end{equation}
such that it represents the error in the ephemeris due to uncertainty
in the masses of MBAs included in the ephemeris (first row), and due
to the entirety of the masses of MBAs omitted from the ephemeris
(second row).

In the third line, we execute the integral over $b,$ and \edited{condense the integrals
over the impulse direction $(\theta,\phi)$ into the angle-averaged
geometry factors $\langle \hat b^2_\gamma\rangle$ of encounters as
a function of $v$.  These will be derived from the actual MBA orbit
distributions.} It must be true that
\begin{equation}
  \left\langle \hat b_r^2 \right\rangle
  + \left\langle \hat b_\phi^2 \right\rangle
  + \left\langle \hat b_z^2 \right\rangle = 1,
\end{equation}
but equality of the three is not required.

@@

\subsection{Numerical results}
\eqq{eq:nI2} reduces the calculation to a double integral over the
deflector's $H$ and the tracer-deflector relative velocity $v.$  The
constituents of this calculation are estimated as follows.

@@

\subsubsection{Asteroid properties}
\label{sec:mbaprops}
We \edited{begin with} the $H$ distribution of potential deflectors,
\edited{since this is directly observable.}.
Figure~\ref{fig:counts} plots our assumptions for $dN/dH.$  The MPC
catalog is likely close to complete in all three regions for $H\le17,$
so we will use the MPC counts directly in this regime.  Note that the
three regions have similar shapes of $dN/dH.$  For $H>17,$ we adopt
the functional form for the cumulative MBA counts given by
initial LSST predictions \citep{lsstbook}:
\begin{equation}
  N(<H) \propto \frac{10^{0.43(H-15.7)}}{10^{0.18(H-15.7)} +
    10^{-0.18(H-15.7)}},
\label{eq:lsstmba}
\end{equation}
normalizing this curve to the observed $H<17$ counts for each region.
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{mbacounts.pdf}
  \caption{Differential counts of MBAs cataloged by the MPC are
    plotted vs absolute magnitude $H$  
    for the Inner, Middle, Outer, and ``Other'' regions of
    the main belt, and for their total.  We assume that they are
    complete for $H\le17,$ and extrapolate to fainter sources using
    the functional form from \eqq{eq:lsstmba} shown as the dotted
    curve. The Inner and Outer regions are slightly rescaled to help
    visual comparison of the curves' shapes in the three MBA regions.}
  \label{fig:counts}
\end{figure}

\edited{Next we require a mean mass of MBAs with a given $H,$ ignoring the fact
  that there will be a substantial spread of $M$ vs $H$.  While one can
  calculate a mass under assumptions of a Lambertian spherical body, a known
  geometric albedo, and a known mean density, each of these assumptions would be
  inaccurate and based on a limited set of well-measured MBAs.  It is more
  effective to examine $M(H)$ directly from the slightly-less-limited set of
  MBAs with usefully constrained masses.  If albedo, density, and shape vary
  weakly with $H$, we will expect a relation of the form
  \begin{equation}
        M(H) = M_{10} 10^{-0.6(H-10)}.
        \label{m10}
      \end{equation}
The distribution of $M_{10}$ values for MBAs with masses found in the literature
is presented in Figure~1 and Section~2.1 of \citet{negin}.  We adopt their
selection of a ``high-mass-to-light'' value of
$M_{10}=2\times10^{17}\,\text{kg}=10^{-13} M_\odot$ and a ``low-$M/L$'' value of
$M_{10}=4\times10^{16}\,\text{kg}=2\times10^{-14} M_\odot$ as providing a rough
estimate of the range that the true mean $M_{10}$ might take.}
%       We crudely approximate the population as having a geometric albedo of 0.25,
% density of 2500~kg~m$^{-3},$ and spherical shapes, which leads to the
% conversion
% \begin{equation}
%   M(H) = 1.2\times10^{-17} M_\odot \times 10^{-0.6(H-15)}
% \end{equation}

Our integral for the uncertainty due to unmodelled MBA gravitation
requires the RMS error $\sigma_M(H)$ between the true mass $M$ and the
mass used (if any) in the ephemeris model.
The two largest asteroids, Ceres and Vesta, have masses determined to
high precision by the \textit{Dawn} spacecraft \citep{Ceres,Vesta}.  Thus despite holding
half of the $\approx 10^{-9} M_\odot$ mass of the asteroid belt, their $\sigma_M$
values\footnote{More precisely, the uncertainties in their $GM$ values
  relative to $GM_\odot$.} are 
$\le3\times10^{-15} M_\odot$ and
% Konilov papers: 1.2e-5 km^3/s^2 for Vesta, 4e-4 for Ceres
this unknown part of their contribution to other MBA's motion is 
unimportant.  Aside from a handful of other asteroids in
well-characterized binary systems or visited by spacecraft, most other knowledge of MBA masses
has and will come from mutual gravitational encounters.   An
individual asteroid's mass can be estimated from its measured effect  on the
few MBAs on which it imparts the largest impulses.  The accuracy
$\sigma_M$ of this determination will vary, depending upon the
geometry  and timing of
each deflector MBA's encounters, and the measurement errors on its tracers.  These
factors are independent of the mass of the deflector.

To give some context, a
comprehensive investigation of MBA masses from mutual scattering
events by \citet{goffin} reports typical $\sigma_M\approx
5\times10^{-13}$ for 30 MBAs (although the list is biased toward
better-measured events). But we can expect substantial improvement
in the LSST era when a millions of potential tracer asteroids attain
mas-level astrometry. \citet{negin} use a simulated LSST sequence of
observations to estimate the impulse (and mass) uncertainties from
analyzing LSST's data on mutual encounters of the known asteroids.
\edited{ They find
that the median $\sigma_M$ for MBAs depends substantially on the quality of
pre-LSST astrometry of tracer bodies, varying from $10^{-13.5} M_\odot$ if no
pre-LSST data are used, to $10^{-14.1} M_\odot$ if pre-LSST data fully determine
the tracer state vector at the start of LSST (see their Figure~4 and Section~5).  These values assume that the 20
most informative tracers of a given target asteroid are combined to find the
best estimate of the deflector mass.  We will adopt a value of
$\sigma_M=10^{-14} M_\odot$  as the nominal situation post-LSST. This is on the
optimistic side of the value of pre-LSST data, but also ignores information that
might be gained from mutual encounters with the $>10^7$ new MBAs that LSST is
expected to discover and have available as tracer particles.}

% We will
% assign a value $\sigma_M$ to the RMS uncertainty in MBA mass
% attainable from following a few individual tracers, and calculate
% the Brownian-motion noise as a function of $\sigma_M.$

% \begin{figure}
%   \plottwo{sigmaI.pdf}{sigmaM.pdf}
%   \caption{Forecasts of uncertainties derived from LSST tracking of the known MBAs.  The left-hand plot shows the distribution of (log) uncertainties $\sigma_I$ on the amplitude of the impulse applied to the tracer MBA, incorporating forecasted LSST observations and marginalizing over the tracer's initial state vector.  The right-hand side shows the distribution of the uncertainties on the mass, $\sigma_M=(bv/2)\sigma_I,$ for deflector MBAs expected from LSST.  The histograms show the results from using only the single tracer yielding the lowest $\sigma_M$ for that deflector; combining the 10 best-$\sigma_M$ tracers; or combining all impulses at $b<0.03$~AU.  The RMS value of $\sigma_M$ over the deflector population is given in the legend.  Masses are in solar units, and velocities in units of the circular velocity of the deflector.}
%   \label{fig:sigmaI}
% \end{figure}
